\MakeUppercase{The Structure of $k$-Planar Graphs}

### 2020-03-25 k-planar.tex
Addressed X versus V(X) issue

Apex-minor-free graphs, addressed by \cref{ApexMinorFree}, are the largest minor-closed class that admit bounded layered partitions~\citep{dujmovic.joret.ea:planar}. However, the family of $k$-planar graphs is not minor-closed.  A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar. \snote{DCG Reviewer \#1 writes: \emph{Isn't this trivial? A grid with all
diagonals will do.}  Are they correct? If so, what to do?} Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors. Our results for $k$-planar graphs are the first of this type for a non-minor-closed class.

@@

\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{thm:ps} and \cref{PlanarPartition}.\ref{subthm:three-tree}

We make use of the following lemma due to \citet{pilipczuk.siebertz:polynomial-soda}.

\begin{lem}[\citep{pilipczuk.siebertz:polynomial-soda}]
\label{p-centered-treewidth}
Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\leq  (t+1)\binom{p+t}{t}$.
\end{lem}

@@

 First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$. Since $G[X]$ is connected, $G[X]$ contains a path from $v$ to $w$.  By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$. Therefore, $|\{\beta(v'):v'\in X\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.

 Otherwise, $X\subseteq L_{i},\ldots,L_{i+s}$ for some $s<p$.  Let $H':=H[\{x\in V(H):B_x\cap X\neq\emptyset]$.  If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in X\}|> p$ so $|\{\phi(v):v\in X\}|> p$ and we are done.  Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
 For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$. Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$. Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
 \end{proof}

@@

\begin{cor}
\label{pck}
For every $k$-planar graph $G$,
%for some integer $t\leq \binom{k+4}{3}-1$,
%$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
\[ \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{k+4}{3} \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
%\note{DW}{I previously wrote ``for some integer $t\leq \binom{k+4}{3}-1$'' thinking that $\binom{p+t}{t}$ might not be monotone with $t$, but it is because of the $+t$.}
%Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
\end{cor}

\begin{cor}
\label{pc1}
For every $1$-planar graph $G$,
\[ \chi_p(G)\le 20 (p+3)(p+2)(p+1)^2\enspace. \]
%120 (p+1) \binom{p+3}{3}
%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
%Every 1-planar graph has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
\end{cor}

\begin{cor}
\label{pcgk}
For every $(g,k)$-planar graph $G$,
%for some integer $t\leq \binom{k+5}{4}-1$,
%$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{k+5}{4}\binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
%O(gk^2)\times p^{O(k^3)}$.
%%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
%Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most
%$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
\end{cor}

>>>>

Apex-minor-free graphs, addressed by \cref{ApexMinorFree}, are the largest minor-closed class that admit bounded layered partitions~\citep{dujmovic.joret.ea:planar}. However, the family of $k$-planar graphs is not minor-closed.  A graph $G'$ obtained from a $k$-planar graph $G$ by edge deletions and edge contractions may or may not be $k$-planar. \snote{DCG Reviewer \#1 writes: \emph{Isn't this trivial? A grid with all
diagonals will do.}  Is the reviewer correct? If so, what to do?} Indeed, \citet{dujmovic.eppstein.ea:structure} construct 1-planar graphs that contain arbitrarily large complete graph minors. Our results for $k$-planar graphs are the first of this type for a non-minor-closed class.

@@

\citet{pilipczuk.siebertz:polynomial-soda} and \citet{debski.felsner.ea:improved} use \cref{thm:ps} and \cref{PlanarPartition}.\ref{subthm:three-tree}, respectively, to show that, when $G$ is planar, or of bounded genus, $\chi_p(G)$ is polynomial in $p$.  We make use of the following lemma due to \citet[Lemma~15]{pilipczuk.siebertz:polynomial-arxiv}.

\begin{lem}[\citep{pilipczuk.siebertz:polynomial-soda}]
\label{p-centered-treewidth}
  Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\leq \binom{p+t}{t}$.
\end{lem}

@@

 First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$. Since $X$ is connected, $X$ contains a path from $v$ to $w$.  By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$. Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.

 Otherwise, $V(X)\subseteq L_{i},\ldots,L_{i+s}$ for some $s<p$.  Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.  If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.  Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
 For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$. Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$. Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
 \end{proof}

@@

\begin{cor}
\label{pck}
For every $k$-planar graph $G$,
%for some integer $t\leq \binom{k+4}{3}-1$,
%$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
\[ \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
%\note{DW}{I previously wrote ``for some integer $t\leq \binom{k+4}{3}-1$'' thinking that $\binom{p+t}{t}$ might not be monotone with $t$, but it is because of the $+t$.}
%Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
\end{cor}

\begin{cor}
\label{pc1}
For every $1$-planar graph $G$,
\[ \chi_p(G)\le 5 (p+3)(p+2)(p+1)^2\enspace. \]
%120 (p+1) \binom{p+3}{3}
%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
%Every 1-planar graph has a layered $H$-partition of layered width at most 30 where $H$ is planar and has treewidth at most 3.
\end{cor}

\begin{cor}
\label{pcgk}
For every $(g,k)$-planar graph $G$,
%for some integer $t\leq \binom{k+5}{4}-1$,
%$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
%O(gk^2)\times p^{O(k^3)}$.
%%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$.
%Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most
%$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
\end{cor}
